---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/hypertidy/laridae/workflows/R-CMD-check/badge.svg)](https://github.com/hypertidy/laridae/actions)
<!-- badges: end -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%", 
  fig.width=12, fig.height=8
)
```

`laridae` came out of a need for constrained triangulation for a topology-in-R project. That effort has moved on somewhat, proving the case by using `RTriangle` and then bedding down the normalization model in the `hypertidy/silicate` package. 

RTriangle is really fast, but it's not as fast as CGAL. CGAL can also be used to update a triangulation, which means (I think) that we could build an unconstrained triangulation from all the coordinates, and then add in any segments, even unclosed linear paths. At any rate, being able to update a mesh has a lot of applications, especially for neighbouring shapes, and for on-demand (extent or zoom dependent level of detail) tasks. 

We can now hit this more easily thanks to [cgal4h](https://CRAN.R-project.org/package=cgal4h) thanks 
to Ahmadou Dicko. 

The interest (which is miniscule ...) in constrained triangulations is discussed here along with the overall landscape in R. 

https://github.com/r-spatial/discuss/issues/6

## Installation



## Triangulation


Triangulate with CGAL via [laridae](https://github.com/hypertidy/laridae).  The function `tri_xy` performs an exact Delaunay triangulation on all vertices, returning a triplet-index for each triangle (zero-based in CGAL). 


The variants `tri_xy1()` and `tri_xy2()` work slightly differently illustrating CGAL usage in C++ (thanks to Mark Padgham). 'xy1' order is not trivial ... but we ignore that for now. 

```{r triangle}
library(laridae)
x <- c(0, 0, 1)
y <- c(0, 1, 1)
plot(x, y, pch = "+")
(idx0 <- tri_xy(x, y))
(idx1 <- tri_xy1(x, y))
(idx2 <- tri_xy2(x, y))

polygon(cbind(x, y)[idx0, ])

x    <- c(2.3,3.0,7.0,1.0,3.0,8.0)
y    <- c(2.3,3.0,2.0,5.0,8.0,9.0)
idx <- tri_xy(x, y)
plot(x, y, pch = "+", cex = 1.5)
polygon(cbind(x, y)[rbind(matrix(idx, 3L), NA), ], lty = 3)

```

Some timings, to show we aren't wildly off-base and that CGAL wins for raw unconstrained Delaunay triangulation.  

```{r}
set.seed(90)
x <- rnorm(1e3, sd = 4)
y <- rnorm(1e3, sd = 2)
library(laridae)

# plot a matrix xy as points
# and add the triangulation indexed
# by structural triplet row-identifiers
poly_index <- function(xy, index, ...) {
  plot(xy, ...)
  ## assume index is 0,1,2,0,1,2,0,1,...
  ii <- c(rbind(matrix(index, nrow = 3), NA_integer_))
  ## have forgetten why polypath fails, so just use polygon
  polygon(xy[ii, 1], xy[ii, 2])
}


ps <- RTriangle::pslg(P = cbind(x, y))

microbenchmark::microbenchmark(
  ind_t <- tri_xy(x, y), 
  ind_t1 <- tri_xy1(x, y), 
  ind_t2 <- tri_xy2(x, y), 
  RT <- RTriangle::triangulate(ps)
)
length(ind_t)
length(ind_t1)
length(ind_t2)
length(RT$T)


p <- par(mfrow = c(2, 2), mar = rep(0, 4))
poly_index(cbind(x, y), ind_t, pch = ".")
## can't work as order is not aligned, but still fun
poly_index(cbind(x, y), ind_t1, pch = ".")  
poly_index(cbind(x, y), ind_t2, pch = ".")
plot(RT)
par(p)



```

## Constrained triangulation

Currently laridae only has "reporting" of the result. I can't yet see how to 

* get the vertex pool (it may have expanded given mesh properties, overlapping segments, etc.)
* get the triangle index

The only other implementation in R is in RTriangle, so we use that for comparison. 


```{r mesh-input, echo=TRUE}
sfx <- silicate::inlandwaters
sc <- silicate::SC(sfx)
X <- sc$vertex$x_
Y <- sc$vertex$y_
i0 <- match(sc$edge$.vx0, sc$vertex$vertex_)
i1 <- match(sc$edge$.vx1, sc$vertex$vertex_)

system.time(insert_constraint(X, Y, i0 , i1))
system.time(segment_constraint(sc))

system.time(insert_mesh(X, Y, i0 , i1))

## compare RTriangle, it's fast if we don't include pslg() time
ps <- RTriangle::pslg(cbind(X, Y), S = cbind(i0, i1))
system.time({
  tr <- RTriangle::triangulate(ps, D = TRUE)
})


plot(tr$P, pch= ".")
segments(tr$P[tr$E[,1],1], tr$P[tr$E[,1],2], 
         tr$P[tr$E[,2],1], tr$P[tr$E[,2],2])

str(tr)
```



## History

Was originally called `cgalgris`. 



## Code of Conduct
  
Please note that the laridae project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/1/0/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
