---
output: github_document
editor_options: 
  chunk_output_type: console
---


[![Travis-CI Build Status](https://travis-ci.org/hypertidy/laridae.svg?branch=master)](https://travis-ci.org/hypertidy/laridae)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/hypertidy/laridae?branch=master&svg=true)](https://ci.appveyor.com/project/hypertidy/laridae)


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

`laridae` came out of a need for constrained triangulation for a topology-in-R project. That effort has moved on somewhat, proving the case by using `RTriangle` and then bedding down the normalization model in the `hypertidy/silicate` package. 

RTriangle is really fast, but it's not as fast as CGAL. CGAL can also be used to update a triangulation, which means (I think) that we could build an unconstrained triangulation from all the coordinates, and then add in any segments, even unclosed linear paths. At any rate, being able to update a mesh has a lot of applications, especially for neighbouring shapes, and for on-demand (extent or zoom dependent level of detail) tasks. 

The interest in constrained triangulations is discussed here along with the overall landscape in R. 

https://github.com/r-spatial/discuss/issues/6

## Installation

Dev-only for now

### Linux

Ubuntu/Debian

```bash
apt install libcgal-dev
apt install libcgal-demo
apt install cmake g++
```

## Other OS ...

## And then

Make sure to run this when your defs change, also  when the system has been updated ? 

```R
tools::package_native_routine_registration_skeleton("../laridae", "src/init.c",character_only = FALSE)
```
WIP


## Triangulation


Triangulate with CGAL via [laridae](https://github.com/hypertidy/laridae).  The function `tri_xy` performs an exact Delaunay triangulation on all vertices, returning a triplet-index for each triangle (zero-based in CGAL). 

Some timings, to show we aren't wildly off-base and that CGAL wins for raw unconstrained Delaunay triangulation. 

```{r}
#x    <- c(2.3,3.0,7.0,1.0,3.0,8.0)
#y    <- c(2.3,3.0,2.0,5.0,8.0,9.0)

set.seed(90)
x <- rnorm(1e3, sd = 4)
y <- rnorm(1e3, sd = 2)
#x <- c(0, 0, 1, 1)
#y <- c(0, 1, 1, 0)
library(laridae)

# plot a matrix xy as points
# and add the triangulation indexed
# by structural triplet row-identifiers
poly_index <- function(xy, index, ...) {
  plot(xy, ...)
  ## assume index is 0,1,2,0,1,2,0,1,...
  ii <- c(rbind(matrix(index, nrow = 3), NA_integer_))
  ## have forgetten why polypath fails, so just use polygon
  polygon(xy[ii, 1], xy[ii, 2])
}


library(dplyr)
library(tibble)

xy <- cbind(x, y)
system.time({
  ind_t <- tri_xy(xy[,1], xy[,2]) + 1
})
system.time({
  ind_t1 <- tri_xy1(xy[,1], xy[,2]) + 1
})
system.time({
  ind_t2 <- tri_xy2(xy[,1], xy[,2]) + 1
})

length(ind_t)
length(ind_t1)
length(ind_t2)


ps <- RTriangle::pslg(P = xy)
system.time({
  ind_T <- c(t(RTriangle::triangulate(ps)$T))
})
length(ind_T)

p <- par(mfrow = c(2, 2), mar = rep(0, 4))
poly_index(xy, ind_t, pch = ".")
## can't work as order is not aligned, but still fun
poly_index(xy, ind_t1, pch = ".")  
poly_index(xy, ind_t2, pch = ".")
poly_index(xy, ind_T, pch = ".")
par(p)


## other comparisons
library(deldir)
system.time(dl <- deldir::deldir(x, y))
plot(dl)
library(geometry)
system.time(gm <- geometry::delaunayn(xy))
poly_index(xy, c(t(gm)))

## sf comparison
library(dplyr)
library(sf)
d <- st_as_sf(tibble::as_tibble(xy) %>% mutate(a = row_number()), coords = c("x", "y"))
## timing is unfair as sf must be decomposed and recomposed
## and every triangle has four coordinates, no sharing allowed
## and probably sfdct is slow ..
library(sfdct)
## this doesn't do anything, same as rgl::triangulate must
## have edge inputs
##system.time(sfd <- st_triangulate(d))
system.time(dt <- ct_triangulate(d))
plot(dt, col = "transparent", border = "black")
```

## Constrained triangulation

There are various ways to do this, but the lowest overhead to start with is to pass in unique vertices and segments pairs in a list. 


```{r}
library(laridae)
library(silicate)


library(dplyr)
prepare_sf_ct0 <- function(x) {
  tabs <- silicate::SC0(x)
  segs <- do.call(rbind, lapply(tabs$object$topology_, as.matrix))
  segs1 <- cbind(.vx0 = pmin(segs[,1], segs[,2]), .vx1 = pmax(segs[,1], segs[,2]))
  print(nrow(segs1))
  
  segs1 <- tibble::as_tibble(segs1) %>% distinct()
  print(nrow(segs1))
  
    list(x = tabs$vertex$x_, y = tabs$vertex$y_, 
         segs = split(t(as.matrix(segs1)), rep(seq_len(nrow(segs1)), each = 2L)))
}
prepare_sf_ct <- function(x) {
  tabs <- silicate::SC(x)
  
  segment <-  tibble::tibble(vertex_ = c(t(as.matrix(sc_edge(tabs) %>% dplyr::select(.vx0, .vx1))))) %>%
  inner_join(tabs$vertex %>% mutate(vertex = row_number() - 1)) %>% mutate(segment = (row_number() + 1) %/% 2)
  segs <- split(segment$vertex, segment$segment)

  list(x = tabs$vertex$x_, y = tabs$vertex$y_, segs = distinct_uord_segments(segs))
}

distinct_uord_segments <- function(segs) {
  x <- dplyr::distinct(tibble::as_tibble(do.call(rbind, segs)))
  usort <- do.call(rbind, lapply(segs, sort))
  bad <- duplicated(usort)
  x <- x[!bad, ]
  lapply(split(x, seq_len(nrow(x))), unlist)
}

st_line_from_segment <- function(segs, coords) {
  sf::st_sfc(lapply(segs, function(a) sf::st_linestring(coords[a + 1, ])))
}

#sline <- st_line_from_segment(psf$segs, cbind(psf$x, psf$y))
```

(This is very early, just to prove we run the meshing and succeed). 

```{r}

library(sfdct)
data("minimal_mesh", package = "silicate")


dat <- minimal_mesh
system.time(psf <- prepare_sf_ct(dat))

laridae:::segment_constraint(psf$x, psf$y, psf$segs)



```


## History

Was originally called `cgalgris`. 



## Code of Conduct
  
Please note that the laridae project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/1/0/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
